Constraints on Cosmic-Ray Particle Populations, Magnetic Fields, and Dark Matter from VHE Gamma-Ray Observations of the Coma Cluster of Galaxies with VERITAS

### 2011-07-07 ms.tex
Added a nice table comment to source table and skeleton for Fermi results table
1st draft of CR and B analysis

\newcommand{\vdag}{(v)^\dagger}
\newcommand{\myemail}{nkarlsson@physics.umn.edu}
\newcommand{\expval}[1]{\left\langle #1 \right\rangle}
\newcommand{\RA}[3]{#1$^{\mathrm{h}}$#2$^{\mathrm{m}}$#3$^{\mathrm{s}}$}
\newcommand{\Dec}[3]{#1$^{\circ}$#2\arcmin#3\arcsec}

@@

%\title{VHE Gamma-Ray Observations of the Coma Cluster of Galaxies with VERITAS}
\title{Constraints on Cosmic-Ray Particle Populations, Magnetic Fields, and Dark Matter from VHE Gamma-Ray Observations of the Coma Cluster of Galaxies with VERITAS}
\author{N. Karlsson\altaffilmark{1,2}, M. Pohl\altaffilmark{3,4}}
\affil{(the VERITAS collaboration)} 
%%\email{nkarlsson@physics.umn.edu}
\and
\author{S. Federici\altaffilmark{3,4}, C. Pfrommer\altaffilmark{5,6},
A. Pinzke\altaffilmark{7}}

\altaffiltext{1}{Corresponding author: N. Karlsson, nkarlsson@physics.umn.edu}
\altaffiltext{2}{School of Physics and Astronomy, University of Minnesota, Minneapolis, MN 55455, USA}
\altaffiltext{3}{DESY, Platanenallee 6, 15738 Zeuthen, Germany}
\altaffiltext{4}{ Institut fu\"r Physik und Astronomie, Universit\"at Potsdam, 14476 Potsdam-Golm, Germany}
\altaffiltext{5}{Heidelberg Institute	for Theoretical	Studies, Schloss-Wolfsbrunnenweg 35, D-69118 Heidelberg, Germany}
\altaffiltext{6}{Canadian Institute for Theoretical Astrophysics, University of Toronto, Toronto, ON, M5S 3H8, Canada}
\altaffiltext{7}{University of California, Santa Barbara, CA 93106}

@@

The Coma cluster of galaxies (ACO 1656) is one of the most thouroughly studied clusters over all wavelengths \citep{article:Voges_etal:1999}. Located at a distance of about 100 Mpc \citep[$z=0.023$;][]{article:StrubleRood:1999}, it is one of the closest and also one of the most massive clusters \citep[$M \sim 10^{15}M_{\odot}$;][]{article:Smith:1983, article:Kubo_etal:2008}. It hosts both a giant radio halo \citep{article:Giovannini_etal:1993,article:Thierbach_etal:2003} and peripheral radio relic, which appears connected to the radio halo with a ``diffuse'' bridge \citep[see discussion in][]{article:BrownRudnick:2010}. It has been suggested that the relic is an infall shock. Extended soft X-ray (SXR) emission is evident from the ROSAT all-sky survey in the 0.1 to 2.4 keV band \citep{article:BrielHenryBohringer:1992} and observations with XMM-Newton \citep{article:Briel_etal:2001} revealed substructure in the X-ray halo. As such it is a natural candidate for gamma-ray observations. In this article, results from the VERITAS observations of the Coma cluster of galaxies are reported, together with the resulting constraints on cosmic-ray particle populations, magnetic fields, and dark matter in the cluster.

@@

\section{Constrainsts on Magnetic Fields}

@@

@@

\begin{figure*}
\begin{center}
\scalebox{1.0}{\plotone{f5.eps}}
\end{center}
\caption{Cosmic-ray pressure, $X_{\mathrm{CR}}$, as a function of radial distance from the Coma cluster core.}
\label{fig:xcr}
\end{figure*}

@@

>>>>

\def\del#1{{}}
%\def\del#1{{\bf (DELETED TEXT)}}
\def\C#1{#1}
%\def\C#1{{\bf #1}}

\newcommand{\vdag}{(v)^\dagger}
\newcommand{\myemail}{nkarlsson@physics.umn.edu}
\newcommand{\expval}[1]{\left\langle #1 \right\rangle}
\newcommand{\RA}[3]{#1$^{\mathrm{h}}$#2$^{\mathrm{m}}$#3$^{\mathrm{s}}$}
\newcommand{\Dec}[3]{#1$^{\circ}$#2\arcmin#3\arcsec}
\newcommand{\rmn}{\mathrm}
\newcommand{\CR}{\mathrm{CR}}
\newcommand{\dps}{\displaystyle}
\newcommand{\eps}{\varepsilon}
\newcommand{\bra}{\langle}
\newcommand{\ket}{\rangle}

@@

%\title{VHE Gamma-Ray Observations of the Coma Cluster of Galaxies with VERITAS}
\title{Constraints on Cosmic-Ray Particle Populations, Magnetic Fields, and Dark Matter from VHE Gamma-Ray Observations of the Coma Cluster of Galaxies with VERITAS}
\author{N. Karlsson\altaffilmark{1,2}, M. Pohl\altaffilmark{3,4}}
\affil{(the VERITAS collaboration)} 
%%\email{nkarlsson@physics.umn.edu}
\and
\author{S. Federici\altaffilmark{3,4}, C. Pfrommer\altaffilmark{1,5,6},
A. Pinzke\altaffilmark{7}}

\altaffiltext{1}{Corresponding authors: N. Karlsson, nkarlsson@physics.umn.edu \& C. Pfrommer, christoph.pfrommer@h-its.org}
\altaffiltext{2}{School of Physics and Astronomy, University of Minnesota, Minneapolis, MN 55455, USA}
\altaffiltext{3}{DESY, Platanenallee 6, 15738 Zeuthen, Germany}
\altaffiltext{4}{ Institut fu\"r Physik und Astronomie, Universit\"at Potsdam, 14476 Potsdam-Golm, Germany}
\altaffiltext{5}{Heidelberg Institute	for Theoretical	Studies, Schloss-Wolfsbrunnenweg 35, D-69118 Heidelberg, Germany}
\altaffiltext{6}{Canadian Institute for Theoretical Astrophysics, University of Toronto, Toronto, ON, M5S 3H8, Canada}
\altaffiltext{7}{University of California, Santa Barbara, CA 93106}

@@

The Coma cluster of galaxies (ACO 1656) is one of the most thouroughly studied clusters over all wavelengths \citep{article:Voges_etal:1999}. Located at a distance of about 100 Mpc \citep[$z=0.023$;][]{article:StrubleRood:1999}, it is one of the closest and also one of the most massive clusters \citep[$M \sim 10^{15}M_{\odot}$;][]{article:Smith:1983, article:Kubo_etal:2008}. It hosts both a giant radio halo \citep{article:Giovannini_etal:1993,article:Thierbach_etal:2003} and peripheral radio relic, which appears connected to the radio halo with a ``diffuse'' bridge \citep[see discussion in][]{article:BrownRudnick:2010}. It has been suggested that the relic is an infall shock \citep{1998A&A...332..395E} and succesively demonstrated by cosmological simulations that model the non-thermal emission processes \citep{2008MNRAS.385.1211P,2008MNRAS.385.1242P,2009MNRAS.393.1073B, 2011ApJ...735...96S}. Extended soft X-ray (SXR) emission is evident from the ROSAT all-sky survey in the 0.1 to 2.4 keV band \citep{article:BrielHenryBohringer:1992} and observations with XMM-Newton \citep{article:Briel_etal:2001} revealed substructure in the X-ray halo. As such it is a natural candidate for gamma-ray observations. In this article, results from the VERITAS observations of the Coma cluster of galaxies are reported, together with the resulting constraints on cosmic-ray particle populations, magnetic fields, and dark matter in the cluster. Throughout the analysis, we use a present day Hubble constant of $H_0 = 100 h \mbox{ km s}^{-1} \mbox{ Mpc}^{-1}$ with $h=0.7$.

@@

We decided to adopt a multifaceted approach to constrain the CR-to-thermal
pressure distribution in Coma with these upper limits. That includes (1) an
analytic model derived from cosmological hydrodynamical simulations of the
formation of galaxy clusters, (2) a simplified multi-frequency analytical
approach that assumes a constant CR-to-thermal energy density and a power-law
spectrum in momentum, and (3) an approach that uses the observed luminosity of
the giant radio halo in Coma to place a lower limit on the expected $\gamma$-ray
flux in the hadronic model -- independent of the magnetic field distribution.
This translates into a minimum CR pressure which, if challenged by tight
gamma-ray limits/detections, enables scrutinizing the hadronic interaction model
of the formation of giant radio halos.  Modelling the CR distribution through
different techniques enable to bracket our ignorance about the underlying plasma
physics that shape the CR distribution and hence reflect the Bayesian priors
that are imposed on the modelling \citep[see][for a
discussion]{2011arXiv1105.3240P}.

\subsection{Simulation-based approach}

We adopt the universal spectral and spatial gamma-ray model developed by
\citet{2010MNRAS.409..449P} to estimate the emission from decaying neutral pions
that dominates over the IC emission from primary and secondary electrons above
100~MeV in clusters. Given a density profile as, e.g., inferred by cosmological
simulations or X-ray observations, the analytic approach models the CR
distribution and the associated radiative emission processes from radio to the
gamma-ray band.  This formalism was derived from high resolution simulations of
clusters of galaxies that included radiative hydrodynamics, star formation,
supernova feedback, and followed CR physics using a novel formulation that trace
the most important injection and loss processes self-consistently while
accounting for the CR pressure in the equation of motion
\citep{2006MNRAS.367..113P,2007A&A...473...41E,2008A&A...481...33J}.  The
results are in line with earlier numerical results on some of the overall
characteristics of the CR distribution and the associated radiative emission
processes \citep{2000A&A...362..151D,2001ApJ...559...59M,2003MNRAS.342.1009M,
  2007MNRAS.378..385P,2008MNRAS.385.1211P,2008MNRAS.385.1242P}, but for the
first time is able to track the spectral evolution as a function of space while
accounting for the non-linear feedback of such a population on the hydrodynamics
in a high-resolution cosmological simulation of the formation of galaxy
clusters.

We note that the overall normalization of the CR and gamma-ray distribution
scales with the maximum acceleration efficiency at structure formation shock
waves. Following recent observations at supernova remnants
\cite{2009Sci...325..719H} as well as theoretical studies
\cite{2005ApJ...620...44K}, we adopt an optimistic but nevertheless realistic
value of this parameter and assume that 50\% of the dissipated energy at strong
shocks is injected into CRs while this efficiency rapidly decreases for weaker
shocks \cite{2007A&A...473...41E}. Since the vast majority of internal formation
shocks (merger and flow shocks) are weak shocks with Mach numbers $M\lesssim3$
\citep[e.g.,][]{2003ApJ...593..599R,2006MNRAS.367..113P} they do not
significantly contribute to the CR population in clusters. Instead, strong
shocks during the formation epoch of clusters and strong accretion shocks at the
present time (at the boundary of voids and filaments/super-cluster regions)
dominate the injection of CRs which are adiabatically transported into clusters
thereafter. Hence, these model predictions provide a plausible upper limit for
the CR contribution from structure formation shocks in galaxy clusters which can
be scaled down for a lower effective acceleration efficiency in the strong shock
regime.  On the other hand, other possible CR sources such as AGN and starburst
driven galactic winds have been neglected for simplicity which could in
principle increase the expected gamma-ray yield from clusters by an unknown
factor.

There are additional limitations of the simulated physics in these cosmological
simulations which only considered advective transport of CRs by turbulent gas
motions which produces centrally enhanced profiles.  However, active CR
transport such as CR diffusion and streaming tends to drive the CR radial
profiles towards being flat, with equal CR number density everywhere. While the
CR streaming is then usually larger than typical advection velocities and
becomes comparable or lower than this only for periods with trans- and
supersonic cluster turbulence during a cluster merger. As a consequence a
bimodality of the CR spatial distribution results with merging (relaxed)
clusters showing a centrally concentrated (flat) CR energy density profile
\cite{2011A&A...527A..99E}.  This translates into a bimodality of the expected
diffuse radio and gamma-ray emission of clusters, since more centrally
concentrated CR will find higher target densities for hadronic CR proton
interactions \cite{2011A&A...527A..99E}. As a result of this, relaxed clusters
could have a reduced gamma-ray luminosity by up to a factor of five. The goal of
this and other future work of seeking for gamma-ray emission from clusters is to
constrain the acceleration physics and transport properties of CRs.

We adopt the electron density profile for the Coma cluster that has been
inferred from ROSAT X-ray observations \citep{1992A&A...259L..31B} and model the
spatial and spectral distribution of CRs according to the (conservative) model
by \citet{2010MNRAS.409..449P} that neglects the contribution of cluster
galaxies. Figure ?? shows the expected spectrum of Coma, normalized to the total
flux within the virial radius. Add {\em Fermi} results \ldots

Figure~\ref{fig:xcr} shows the relative CR pressure, $X_\CR = \bra P_\CR
\ket / \bra P_\rmn{th} \ket$ as a function of radial distance from the
Coma cluster center (dashed lines). To compute the CR pressure, we assume a
lower cutoff of the CR distribution of $P = 0.8\,m_\rmn{p} c^2$ as found in
these simulations due to the high cooling rates at low CR energies that
effectively thermalizes the low-energy, trans-relativistic CR population.  The
relative CR pressure rises toward the outer regions due to CR acceleration being
more efficient at the peripheral strong accretion shocks compared to weak
central flow shocks and adiabatic compression of a composite of CRs and thermal
gas disfavours the CR pressure relative to the thermal pressure due to the
softer equation of state of CRs. The weak increase of $X_\CR$ toward the core is
due to the comparably fast thermal cooling of gas which diminishes the thermal
pressure support relative to that in CRs yielding to adiabatic compression of
CRs.

Table~\ref{table:constraints} shows the expected gamma-ray fluxes within the
regions of interest. For the most constraining regions tested (within an
aperture of radius 0.2$^\circ$), those are a factor of 7.2 below the inferred
upper limits (see Table~\ref{table:results} and assuming a spectral index of
$\alpha=2.3$ which matches the simulated one at energies $E=200$~GeV).  Hence we
scale the averaged relative CR pressure, $X_\CR = \bra P_\CR \ket /
\bra P_\rmn{th} \ket$ of our model by that factor, keep the spatial
behavior and obtain an integrated limit of the relative CR pressure of
$X_\CR<0.091$ within 0.2$^\circ$ that translates to a limit within the cluster
virial radius of $X_\CR<0.13$ (solid lines of Figure~\ref{fig:xcr}).

\subsection{Simplified analytical model}

To complement the simulation-inspired analysis and to facilitate comparison with
earlier analytical work, we repeat the data analysis with a simplified
analytical model that assumes a power-law CR spectrum and a constant relative CR
pressure, i.e., we adopt the {\em isobaric model of CRs}. Since a priori, the CR
spectral index is unconstrained, we vary it within plausible ranges of
$2.1<\alpha_\rmn{p}<2.5$. These spectral ranges are needed to explain the radio
spectral index of the giant radio halo of Coma after accounting for the spectral
steepening at frequencies $\nu\sim5~\rmn{GHz}$ due to the Sunyaev-Zel'dovich
effect \citep{2004A&A...413...17P}\footnote[4]{Assuming a magnetic field of
  1~$\mu$G, the CR protons responsible for the GHz radio emitting electrons have
  an energy of $~100$~GeV and are $\sim$ 20 times less energetic than those CR
  protons that are responsible for the 200 GeV $\gamma$-ray emission. This is
  consistent with the concave curvature found in the universal CR spectrum
  \citet{2010MNRAS.409..449P}.}.  Table~\ref{table:constraints} shows the
resulting constraints on $X_\CR$ which range between 0.05 and 1, strongly
depending on the assumed spectral index.

Add {\em Fermi} results, critical spectral index \ldots

\del{
the limit for 0.2 deg and alpha_p = 2.3 is compatible with the simulation based
one, X_CR < 0.1. However, there is currently a discrepancy in our assumptions
about the lower energy cutoff of CRs: in the simulations we use P/mp c = 0.8 (as
simulated) while I set this to 0 in my simplified (isobaric) model. The
discrepancy can be understood as follows:
*) in the isobaric model, we would have to normalize our $X_\CR$ by a factor of 2/3
for an effective spectral index of ~ 2.4 (between GeV and 200 GeV) as inferred
from the simulations (see Fig 1, CP & Ensslin 2004, A&A, 413, 17).
*) In the simulation model, we have an increased pressure by this factor of 3/2
(in comparison to a power law spectrum with alpha_TeV ~ 2.3 down to a cutoff in
the normalized momentum of 0.8) due to the concave curvature of our spectrum.
}

\subsection{Minimum gamma-ray flux}

For clusters that host radio halos we are able to derive a minimum $\gamma$-ray
flux in the hadronic model of CR interactions. The idea is based on the fact
that a stationary distribution of CR electrons loses all its energy to
synchrotron radiation for strong magnetic fields ($B \gg B_\rmn{CMB} \simeq 3.2
\mu$G) so that the ratio of $\gamma$-ray to synchrotron flux becomes independent
of the spatial distribution of CRs and thermal gas \citep{2008MNRAS.385.1242P}.
This can be easily seen by considering the pion decay induced $\gamma$-ray
luminosity $L_\gamma$ and the synchrotron luminosity $L_\nu$ of a steady state
distribution of CR electrons that has been generated by hadronic CR
interactions,
\begin{eqnarray}
L_\gamma &=& A_\gamma \dps\int d V\, n_\CR n_\rmn{gas},\\
\label{eq:Lnu}
   L_\nu &=& A_\nu \dps\int d V\, n_\CR n_\rmn{gas} \frac{\dps
    \eps_B^{(\alpha_\nu+1)/2}}{\dps\eps_\rmn{CMB}+\eps_B}\\
  &\simeq&A_\nu \dps\int d V\, n_\CR n_\rmn{gas}\quad\rmn{for}~\eps_B \gg \eps_\rmn{CMB}.
\end{eqnarray}
Here $A_\gamma$ and $A_\nu$ are dimensional constants that depend on the
hadronic physics of the interaction \citep{2008MNRAS.385.1211P,
  2008MNRAS.385.1242P} and $\alpha_\nu\simeq 1$ is the observed synchrotron
spectral index. Hence we can derive a minimum $\gamma$-ray flux in the hadronic
model
\begin{equation}
  \label{eq:Fmin}
  \mathcal{F}_{\gamma,\rmn{min}} = 
  \frac{\dps A_\gamma}{\dps A_\nu}\frac{\dps L_\nu}{\dps 4\pi D_\rmn{lum}^2},
\end{equation}
where $L_\nu$ is the observed luminosity of the radio mini-halo and
$D_\rmn{lum}$ denotes the luminosity distance to the respective cluster.
Lowering the magnetic field would require an increase in the energy density of
CR electrons to reproduce the observed synchrotron luminosity and thus increase
the associated $\gamma$-ray flux.

To derive a minimum gamma-ray flux that is comparable to the upper limits, we
need to determined the radio flux within the corresponding angular regions. To
this end, we fit the point source subtracted, azimuthally averaged radio halo
profile at 1.38 GHz \citet{1997A&A...321...55D} with a $\beta$-model,
\begin{equation}
S_\nu (r_\bot)= S_0\, 
\left[ 1 + \left( \frac{r_\bot}{r_{\rmn{c}}}\right)^2\right]
^{-3\beta + 1/2},
\label{beta}
\end{equation}
where $S_0 = 1.1 \times 10^{-3}\,\rmn{Jy\,arcmin}^{-2}$, $r_{\rmn{c}} = 450
kpc$, and $\beta = 0.78$.  Within the error bars, this profile is consistent
with 326 MHz-data taken by \citet{2001A&A...369..441G} when scaled with a radio
spectral index of 1.15.

The results for the minimum gamma-ray flux $F_{\rmn{min}}(>E=220~\rmn{GeV}$ and
the minimum relative CR pressure, $X_{\CR,\,\rmn{min}} = F_{\rmn{min}}/F$ are
shown in Table~\ref{table:constraints}. Even in the most constraining case
($\alpha=2.1$, $<0.2^\circ$), these are a factor of about 60 below the upper
limits obtained in this work.

Add {\em Fermi} results\ldots

\section{Constraining the Magnetic Field}
\label{sec:B}

We have seen in the previous Section that we can obtain an absolute lower limit
on the gamma-ray emission in the hadronic model by assuming high magnetic
fields, $B>B_\rmn{CMB}$. Hence, we can turn this around by using our upper limit
on the gamma-ray emission (and by extension on the CR pressure) to infer a lower
limit on the magnetic field that is needed to explain the observed radio
emission which, however, assumes the validity of the hadronic model. Lowering
the gamma-ray constraint will tighten (increase) the magnetic field limit. Once
this gets in conflict with magnetic field measurements by means of other
methods, e.g., Faraday rotation measure (RM)\footnote{Generally, Faraday RM
  of the magnetic field strength by, e.g., background sources through clusters
  are degenerate with the magnetic coherence scale. Only a simultaneous
  deprojection method that self-consistently solves for the magnetic power
  spectrum is in principle able to break this degeneracy, however, require
  high-quality data and picks up further complications when using embedded
  sources within the cluster plasma, e.g., degeneracies with the assumed
  projection angle of the jet \citep{2011A&A...529A..13K}.}, this would
challenge the hadronic model of radio halos.  The method that we use to
constrain the magnetic field inherits a dependence on the assumed spatial
structure that we parametrize by
\begin{equation}
  \label{eq:B}
  B(r) = B_0  \,\left(\frac{n_\rmn{e}(r)}{n_\rmn{e}(0)}\right)^{\alpha_B},
\end{equation}
as suggested by Faraday RM studies and numerical magneto-hydrodynamical (MHD)
simulations \citep[][and references therein]{2010A&A...513A..30B,
  2011arXiv1107.0968B}. In fact, the magnetic field in the Coma cluster is among
the best ones constrained due to its proximity that enables RM observations of
seven radio-sources located at projected distances of 50 to 1500 kpc from the
cluster center. The best fit model obtains $B_0 = 4.7^{+0.7}_{-0.8}\,\mu$G and
$\alpha_B = 0.5^{+0.2}_{-0.1}$ \citep{2010A&A...513A..30B}.  We aim for
constraining the central field strength, $B_0$, which depends on the magnetic
decline, $\alpha_B$, that we vary (within reasonable ranges of
$\Delta\alpha_B=0.2$ as suggested from the Faraday RM studies) to address this
dependence. We proceed as follows:

1) Given a model for the magnetic field, we determine the profile of the
relative CR pressure, $X_\CR$, so that the hadronically produced CR electrons
exactly matches the observed radio halo emission over the entire extent. To this
end, we de-project the fit to the surface brightness profile (using an Abel
integral equation, see Appendix of \citealt{2004A&A...413...17P}) yielding the
radio emissivity,
\begin{equation}
  \label{eq:Coma:radio}
  j_\nu (r) = \frac{S_0}{2\pi\, r_\rmn{c}}\, 
  \frac{6\beta - 1}{\left(1 + r^2/r_\rmn{c}^2\right)^{3 \beta}}\,
  \mathcal{B}\left(\frac{1}{2}, 3\beta\right)
  = j_{\nu,0} \left(1 + r^2/r_\rmn{c}^2\right)^{-3 \beta},
\end{equation}
where $\mathcal{B}$ denotes the beta function. Using ``isobaric CR model'' of
the formalism suggested by \citet{2004A&A...413...17P} to compute the secondary
radio emission and imposing the boundary condition to reproduce the observed
radio emission profile uniquely determines $X_\CR(r)$. One of those realizations
for the relative CR energy density $\tilde{X}_\CR(r) = \eps_\CR (r) /
\eps_\rmn{th} (r)$ and $\tilde{X}_B(r) = \eps_B (r) / \eps_\rmn{th} (r)$ is
shown in Figure 2 (lower left panel) of \citet{2004MNRAS.352...76P} which also
has the interesting property to represent the minimum energy configuration that
is able to reproduce the radio emission in the hadronic model.\footnote{Note
  that a central magnetic field of $3\,\mu$G corresponds in the Coma cluster to
  a relative magnetic energy density of $\tilde{X}_B=0.005$. For a CR population
  in clusters that have been accelerated in structure formation shocks the
  relativistic limit $E_\CR/E_\rmn{th} = 2 \bra P_\CR \ket / \bra P_\rmn{th}
  \ket$ is not applicable since the CR pressure is dominated by the
  trans-relativistic regime. This implies a somewhat harder equation of state
  for the CRs with a larger adiabatic index and yields the relation
  $E_\CR/E_\rmn{th} = 1.6 \bra P_\CR \ket / \bra P_\rmn{th} \ket$.}  The shaded
regions indicate the logarithmic curvature at the minimum, showing that only
modest additional energy investment would be needed to move the system by a
factor two away from this energetically least expensive state which needs not to
be realized in Nature.  The realization of the minimum energy magnetic field
corresponds to a magnetic decline of $\alpha_B\simeq0.3$ and is close to
equipartition with the CR proton pressure (but not that of the resulting
secondary CR electrons).  It is generically true for weak magnetic fields
($B<B_\rmn{CMB}$) in the outer parts of the Coma halo that the product
$X_\CR(r)X_B(r)$ has to increase by a factor of about 100 toward the radio halo
periphery to account for the observed extent.  If we were to adopt a steeper
magnetic decline such as $\alpha_B=0.5$ which produces a flat $X_B(r)$, the
relative CR pressure would have to rise accordingly by a factor of 100.

2) Given this realization for $X_\CR$, we compute the pion decay gamma-ray
surface brightness profile, integrate the flux within a radius of $\{0.13, 0.2,
0.4\}$ degree, and scale the CR profile in order to match the corresponding
VERITAS upper limits.  This scaling factor, $X_{\CR,0}$, depends on the CR
spectral index, $\alpha$, (assuming a power-law CR population for simplicity) and
the radial decline of the magnetic field, $\alpha_B$.

3) In the last step, we scale the secondary radio emission profile linearly
with $X_{\CR,0}$ and solve for $B_0$ subject to the boundary condition to match
the observed synchrotron profile. Note that for $B_0 \gg B_\rmn{CMB}$ and radio
spectral index, $\alpha_\nu$, the solution would be degenerate according to
Eqn.~(\ref{eq:Lnu}).

Table~\ref{table:Bmin} shows the results on the lower limit of the central
magnetic field that range from $B_0 = 0.4$ to $1.2\,\mu$G. There are a few
interesting remarks in order.  (1) The hardest CR spectral indices are most
constraining because of the tightest upper limits.  (2) For a steeper magnetic decline
(larger $\alpha_B$), the CR number density needs to be larger to match the
observed radio emission profiles which would yield a higher gamma-ray flux so
that the upper limits are more constraining. This implies tighter lower limits for $B_0$.
However, the largest choice of $\alpha_B=0.7$, which represents a flux-freezing
magnetic field, is possibly too steep in the outer parts as it neglects the
contributions from the turbulent dynamo in these dynamically active regions with
large amounts of vorticity injection at corrugated structure formation shocks
\citep[as observationally inferred for the first time
by][]{2011ApJ...730...22P}.  (3) Interestingly, in all cases, we use the
0.4$^\circ$ limit that is always most constraining for our choice of magnetic
field configurations.  This is because we need a substantially increasing
relative CR pressure profile to match the observed radio profiles (for a given
magnetic realization); hence that CR realization produces a larger flux within
0.4$^\circ$ in comparison to the ``isobaric CR model'' assumption ($X_\CR =
\rmn{const.}$) for which the 0.2$^\circ$ upper limits are more constraining;
hence the constraints on $B_0$ get better for the larger integration area.
Physically, this can be achieved by considering (active) CR transport such as CR
streaming which produces an increasing $X_\CR$-profile due to the larger
available phase space in the outer cluster regions (or equivalently an
incomplete turbulent pumping of CRs into the central regions).  Required
streaming and turbulent pumping timescales match what is required to explain the
observed radio halo bimodality when carefully considering the plasma physics in
high-beta plasmas \citep{2011A&A...527A..99E}.

As a final word, we note that the $X_\CR$-profiles always obey the energy
condition, i.e., $P_\CR < P_\mathrm{th}$ over the entire range of the radio halo
emission ($< 1$~Mpc) when assuming a central magnetic field of $B_0=3\,\mu$G
that is at the lower 2-$\sigma$ interval bound inferred from Faraday RM
\citep{2010A&A...513A..30B}. Typical values at 1 Mpc range from $X_\CR = 0.02$
($\alpha_B=0.3$, $\alpha=2.1$) to $X_\CR = 1$ ($\alpha_B=0.7$,
$\alpha=2.5$).\footnote{See Figure 3 in \citet{2004MNRAS.352...76P} for the
  entire parameter range assuming minimum energy conditions and
  \citet{2004A&A...413...17P}, Figure 7 for a parametrization as adopted in this
  study.} We caution, however, that the minimum energy condition is violated at
the outer radio halo boundary for the range of minimum magnetic field values
inferred by this study.

@@

Various approaches to model CR population and magnetic field distribution are
complementary (so far just ideas, need to be formulated):
\begin{itemize}
\item simplified ``isobaric CR model'': simple and widely used in the
  literature. A priori, physically not justified, old fashioned, should be
  replaced by the approached below.
\item simulation-based analytical approach: ``first principle approach'' that
  predicts the CR and B distribution for a given set of assumptions. Powerful
  since it only required the density profile as input, but limited in the
  modeled physics (no active CR transport) that appears to be necessary to
  explain the radio halo bimodality.
\item pragmatic approach as employed in Sect.~\ref{sec:B}: model CR and B to
  reproduce observed radio emission profile. Minimum energy conditions (or
  equivalently equipartition arguments) represent a special case within this
  approach. A priori, also physically not justified, but quite powerful as it
  shows what the ``correct'' model has to achieve and can point into the
  direction what the relevant physics has to deliever. Compare CR transport
  vs. simulation inspired models
\end{itemize}

@@

\begin{figure*}
\begin{center}
\scalebox{1.0}{\plotone{f5.eps}}
\end{center}
\caption{CR-to-thermal pressure, $X_{\mathrm{CR}}\bra P_\CR \ket / \bra
  P_\rmn{th} \ket$, as a function of radial distance from the Coma cluster
  center. Model predictions are shown by dashed lines \citep{2010MNRAS.409..449P}
  while these are scaled to match the most constraining VERITAS upper limits
  within 0.2$^\circ$ (solid). We compare differential $X_\CR$-profiles (red) to
  integrated profiles $X_\CR(<R/R_\rmn{vir})$ (blue) which we use to compare to
  the upper limits.}
\label{fig:xcr}
\end{figure*}

@@

\begin{landscape}
\begin{deluxetable}{cccccccccccc}
\tablecolumns{12}
\tablewidth{0pc}
\tablecaption{Constraints on relative CR pressure and magnetic fields for different regions of interest in the Coma cluster of galaxies.}
\tablehead{
\colhead{RoI\tablenotemark{a} [deg]} 
    & \colhead{$F_\rmn{sim}(E)$\tablenotemark{b}} & \colhead{$F_\rmn{sim}(>E)$\tablenotemark{c}} 
    & \multicolumn{3}{c}{analytic model: $X_{\CR,\,\rmn{max}}~[\%]$\tablenotemark{d} } 
    & \multicolumn{3}{c}{$F_{\gamma,\,\rmn{min}}(>E)$\tablenotemark{e}} 
    & \multicolumn{3}{c}{$X_{\CR,\,\rmn{min}}~[\%]$\tablenotemark{f} } \\
& & & \colhead{$\alpha=2.1$} & \colhead{$\alpha=2.3$} & \colhead{$\alpha=2.5$}
    & \colhead{$\alpha=2.1$} & \colhead{$\alpha=2.3$} & \colhead{$\alpha=2.5$}
    & \colhead{$\alpha=2.1$} & \colhead{$\alpha=2.3$} & \colhead{$\alpha=2.5$}
}
\startdata
0.13 & -     & -    & 10   & 23 & 97 & 1.6 & 0.7 & 0.3 & 0.067 & 0.061 & 0.11 \\
0.2  & 0.80  & 2.9  & 4.8  & 10 & 43 & 3.1 & 1.4 & 0.6 & 0.078 & 0.072 & 0.13 \\
0.4  & 1.21  & 4.4  & 6.7  & 15 & 62 & 6.3 & 2.8 & 1.3 & 0.098 & 0.090 & 0.16
\enddata
\tablenotetext{a}{Intrinsic source radius (in the point source case, the radius got replaced by the PSF radius).}

\tablenotetext{b}{Differential gamma-ray flux $E^2 dF/dE$ at $E=220$~GeV in
  units of $10^{-6}$ GeV m$^{-2}$ s$^{-1}$ in the simulation-based analytic
  model by \citet{2010MNRAS.409..449P}. }

\tablenotetext{c}{Integrated gamma-ray flux $F(>E=220~\rmn{GeV})$ in units of
  $10^{-9}$ ph. m$^{-2}$ s$^{-1}$ in the simulation-based analytical model by
  \citet{2010MNRAS.409..449P}.}

\tablenotetext{c}{Constraint on the relative CR pressure, $X_\CR = \bra P_\CR\ket /
  \bra P_\rmn{th}\ket$ that was assumed to be constant throught the cluster; calculated
  according to \citet{2004A&A...413...17P}.}

\tablenotetext{e}{Minimum gamma-ray flux $F_{\rmn{min}}(>E=220~\rmn{GeV}$ in units of
  $10^{-10}$ ph. m$^{-2}$ s$^{-1}$ of the hadronic model described in the text.}

\tablenotetext{f}{Minimum relative CR pressure, $X_{\CR,\,\rmn{min}} = F_{\rmn{min}}/F$ for energies
  above $E=220~\rmn{GeV}$ of the hadronic model described in the text.}
\label{table:constraints}
\end{deluxetable}
\end{landscape}

\begin{deluxetable}{cccc}
\tablecolumns{4}
\tablewidth{0pc}
\tablecaption{Constraints on magnetic fields in the hadronic model of the radio halo of Coma}
\tablehead{
\colhead{$\alpha_B$\tablenotemark{a}} &
\multicolumn{3}{c}{Minimum magnetic field, $B_0~[\mu\rmn{G}]$\tablenotemark{a}}\\
\colhead{} & \colhead{$\alpha=2.1$} & \colhead{$\alpha=2.3$} & \colhead{$\alpha=2.5$}
}
\startdata
0.3 & 0.61 & 0.51 & 0.43 \\
0.5 & 0.82 & 0.68 & 0.58 \\
0.7 & 1.19 & 0.98 & 0.83 
\enddata
\tablenotetext{a}{The parameters of the magnetic field are the magnetic decline,
  $\alpha_B$, and the central field strength, $B_0$, which are defined by $B(r) = B_0
  \,[n_\rmn{e}(r)/n_\rmn{e}(0)]^{\alpha_B}$.}
\label{table:Bmin}
\end{deluxetable}
